Non-kissing and non-crossing complexes \\ for locally gentle algebras

### 2018-07-12 draft/accordionComplexSurfaces.tex
first corrections

\begin{definition}%[\cite{PaluPilaudPlamondon}]
\label{def:blossomingQuiver}
A locally gentle bound quiver~$\bar Q$ is \defn{complete} if any vertex~$v \in Q_0$ is incident to either one ($v$ is a leaf) or four arrows ($v$ is an internal vertex).
The \defn{pruned subquiver} of a quiver $\bar Q$ is the locally gentle quiver obtained by deleting all leaves of~$\bar Q$ (degree one vertices) and their incident arrows.
The \defn{blossoming quiver} of a locally gentle bound quiver~$\bar Q$ is the complete locally gentle bound quiver~$\bar Q\blossom$ whose pruned subquiver is~$\bar Q$.
The vertices of~$Q\blossom_0 \ssm Q_0$ are called \defn{blossom vertices}, and the arrows in~$Q\blossom_1 \ssm Q_1$ are called \defn{blossom arrows}.
\end{definition}

Examples are represented on \fref{fig:quivers}.

@@

\begin{definition}
\label{def:string}
Let~$\bar Q = (Q,I)$ be a locally gentle bound quiver.
A \defn{finite string} in~$\bar Q$ is a word of the form
\(
\rho = \alpha_1^{\varepsilon_1}\alpha_2^{\varepsilon_2}\cdots \alpha_\ell^{\varepsilon_\ell},
\)
where
	\begin{itemize}
	\item $\alpha_i \in Q_1$ and~$\varepsilon_i \in \{-1,1\}$ for all~$i \in [\ell]$,
	\item $t(\alpha_i^{\varepsilon_i}) = s(\alpha_{i+1}^{\varepsilon_{i+1}})$ for all~$i \in [\ell-1]$,
	\item there is no path~$\pi \in I$ such that~$\pi$ or~$\pi^{-1}$ appears as a factor of~$\rho$, and
	\item $\rho$ is reduced, in the sense that no factor~$\alpha\alpha^{-1}$ or~$\alpha^{-1}\alpha$ appears in~$\rho$, for~$\alpha \in Q_1$.
	\end{itemize}
The integer~$\ell$ is called the \defn{length} of the string~$\rho$.
We let~$s(\rho) \eqdef s(\alpha_1^{\varepsilon_1})$ and~$t(\rho) \eqdef t(\alpha_\ell^{\varepsilon_\ell})$ denote the source and target of~$\rho$.
For each vertex~$v \in Q_0$, there is also a \defn{string of length zero}, denoted by~$\varepsilon_v$, that starts and ends at~$v$.
%We denote by~$\strings(\bar Q)$ the set of strings on~$\bar Q$.
%We usually denote strings by~$\rho, \sigma, \tau$.
%We call \defn{negative simple string} a formal word of length zero of the form $-v$, where $v$ is any vertex of $Q_0$. We denote by
%\(
%\strings^\pm_{\ge -1}(\bar Q) \eqdef \strings^\pm(\bar Q) \sqcup \set{-v}{v \in Q_0}
%\)
%the set of \defn{almost positive unoriented strings}.
%Note that~$\varepsilon_v$ and~$-v$ are different almost positive strings.
%\end{enumerate}
\end{definition}

\begin{definition}
 An oriented cycle~$c$ in~$Q$ such that~$c$ and~$c^2$ do not belong to the ideal of relations~$I$ is called \defn{primitive} if it cannot be written as an $n$th power ($n>1$) of a cycle.
\end{definition}

\begin{notation}
If~$c$ is an oriented cycle in~$Q$ such that~$c$ and~$c^2$ do not belong to the ideal of relations~$I$, we write:
\begin{itemize}
\item $c^\infty$ for the infinite word~$c c c\cdots$,
\item ${}^\infty c$ for the infinite word~$\cdots c c c$,
% \item $^{-\infty} c$ for the infinite word~$\cdots c^{-1} c^{-1} c^{-1}$,
% \item $c^{-\infty}$ for the infinite word~$c^{-1} c^{-1} c^{-1}\cdots$, and
\item $^\infty c^\infty$ for the bi-infinite word~$^\infty c c^\infty$.
\end{itemize}
We also define~${}^{-\infty}c \eqdef {}^{\infty}(c^{-1}) = (c^\infty)^{-1}$ and similarly~$c^{-\infty} \eqdef (c^{-1})^\infty = (^\infty c)^{-1}$.
\end{notation}

@@

\begin{notation}
To avoid distinguishing between finite, infinite and bi-infinite words, we denote strings as products~$\rho = \prod_{i < \ell < j} \alpha_\ell$ where~$i < j \in \Z \cup \{\pm \infty\}$ and~$\alpha_\ell \in Q_0$.
\end{notation}

@@

\begin{definition}
A \defn{marked surface}~$\bar\surface \eqdef (\surface, M)$ is an orientable surface~$\surface$ with boundaries, together with a set~$M$ of marked points which can be on the boundary of~$\surface$ or not.
For~$V \subset \surface$,
\begin{itemize}
\item a \defn{$V$-arc} on~$\bar\surface$ is a curve on~$\surface$ connecting two points of~$V$ and whose interior is disjoint from~$M$ and the boundary of~$\surface$.
\item a \defn{$V$-curve} on~$\bar\surface$ is a curve on~$\surface$ which at each end either reaches a point of~$V$ or infinitely circles around and finally reaches a puncture of~$M$, and whose interior is disjoint from~$M$ and the boundary of~$\surface$.
\end{itemize}
As usual, curves and arcs are considered up to homotopy relative to their endpoints in~$\surface \ssm M$, and curves homotopic to a boundary are not allowed.
\end{definition}

@@

Some examples of dual cellular dissections on different surfaces are represented in \fref{fig:dissections}.
%
\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[scale=.7]{dissections}}
	\caption{Some pairs~$(\dissection, \dissection\dual)$ of dual cellular dissections on different surfaces. The boundaries of the surfaces are shaded, and the glue symbols indicate how to identify some edges to get the desired surfaces. The first two surfaces are disks, the third is a sphere without boundary, the fourth is a cylinder, and the fifth is a punctured torus. The dissection~$\dissection$ is in green while its dual dissection~$\dissection\dual$ is~in~red.}
	\label{fig:dissections}
\end{figure}
%
Note that contrarily to the usual conventions, the dual vertex~$f\dual$ of a face~$f$ of~$\dissection$ is not always in the interior of the face~$f$.
More precisely, there are two situations:
\begin{itemize}
\item if a face~$f$ has no edge on the boundary of~$\surface$, its dual vertex~$f\dual$ lies in the interior~of~$f$,
\item if a face~$f$ has an edge on the boundary of~$\surface$, then it has exactly one such edge and its dual vertex~$f\dual$ lies on this edge.
\end{itemize}
In fact, the second point forces the following characterization of the cellular dissections that admit a dual cellular dissection.

@@

In this section, we show that non-kissing and non-crossing complexes actually coincide.
Namely, any pair of dual cellular dissections~$(\dissection, \dissection\dual)$ defines a locally gentle bound quiver~$\bar Q_{\dissection}$ such that the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection}]$.
Conversely, any locally gentle bound quiver~$\bar Q$ gives rise to a surface~$\surface_{\bar Q}$ equipped with a pair of dual cellular dissections~$\dissection_{\bar Q}$ and~${\dissection\dual_{\bar Q}}$ such that the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection\dual_{\bar Q}}]$.
In fact, we show that the accordion, slalom, and non-kissing complexes can all be pictured in a unique way on the surface.
Interestingly, via the correspondence between pairs of dual cellular dissections and locally gentle bound quivers, the duality between the dissections~$\dissection$ and~$\dissection\dual$ translates into the Koszul duality between the locally gentle algebras of~$\bar Q$ and~$\bar Q\koszul$.

@@

The advantage of Definition~\ref{def:surfaceQuiver} is that it automatically endows~$\surface_{\bar Q}$ with two disjoint sets~$V_{\bar Q}$ and~${V\dual_{\bar Q}}$ of marked points and two dual cellular dissections~$\dissection_{\bar Q}$ and~${\dissection\dual_{\bar Q}}$ defined as follows.

\begin{definition}
\label{def:dissectionQuiver}
The surface~$\surface_{\bar Q}$ is endowed with
\begin{itemize}
\item the set~$V_{\bar Q}$ of points~$v(\alpha)$ for~$\alpha \in Q_1\blossom$ after the identifications given by~(ii),
\item the $V_{\bar Q}$-dissection~$\dissection_{\bar Q}$ given by all sides~$\Enrs{\alpha}$ and~$\Enrt{\alpha}$ for~$\alpha \in Q_1\blossom$ after the identifications given by~(ii).
\end{itemize}
The set~${V\dual_{\bar Q}}$ and the ${V\dual_{\bar Q}}$-dissection~${\dissection\dual_{\bar Q}}$ are defined dually.
\end{definition}

\begin{proposition}
\label{prop:dissectionsAreCellular}
Let~$\bar Q$ be a locally gentle bound quiver.
Then the dissections~$\dissection_{\bar Q}$ and~$\dissection\dual_{\bar Q}$ are cellular and dual to each other.
\end{proposition}
\begin{proof}
In the construction of~$\surface_{\bar Q}$, if two arrows~$\alpha$ and~$\beta$ of~$\bar Q \blossom$ are such that~${t(\alpha) = s(\beta)}$ and~${\alpha\beta \in I\blossom}$, then the vertices~$v(\alpha)$ and~$v(\beta)$ are identified on the surface.
Thus, any two arrows of a given path contribute the same vertex of~$V_{\bar Q}$, so~$V_{\bar Q}$ is in bijection with the set of maximal paths in~$\bar Q\blossom$.

@@

\begin{remark}
\label{rem:propertiesSurface}
The following observations are useful for the computation of examples.
\begin{enumerate}[(i)]
\item The set~$V_{\bar Q}$ has one vertex for each straight walk in~$\bar Q$ (equivalently, for each maximal path in~$\bar Q$).
      Finite straight walks yield vertices on the boundary, while infinite cyclic straight walks in~$\bar Q$ and in~$\bar Q\koszul$ yield punctures.
\item \label{item:arcs}
      The dissection~$\dissection_{\bar Q}$ has one edge for each vertex~$v \in \bar Q_0$, obtained by concatenation of the sides~$\Enrt{\alpha} = \Enrs{\beta}$ and~$\Enrt{\alpha'} = \Enrs{\beta'}$ where~$v = t(\alpha) = s(\beta) = t(\alpha') = s(\beta')$, $\alpha\beta \notin I$ and~$\alpha'\beta' \notin I$. We denote by~$\edgeof(v)$ the edge of~$\dissection_{\bar Q}$ corresponding to~$v$.
\item The dissection~$\dissection_{\bar Q}$ has one $\ell$-cell for each straight walk of length~$\ell$ in~$\bar Q\koszul$.
\item A similar remark holds dually for~${V\dual_{\bar Q}}$ and~${\dissection\dual_{\bar Q}}$.
\item The number~$p$ of punctures of~$\surface_{\bar Q}$ is the number of infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$.
\item \label{item:boundary}
      The number~$b$ of boundary components of~$\surface_{\bar Q}$ can be computed as follows.
      There are two natural perfect matchings whose vertices are the blossom vertices of~$\bar Q$: one is obtained by joining the endpoints of each finite straight walk of~$\bar Q$, and the other is obtained similarly from~$\bar Q \koszul$.
      Let~$G$ be the superposition of these two perfect matchings.
      Then the number~$b$ of boundary components of~$\surface_{\bar Q}$ is the number of connected components of~$G$.
\item The genus of the surface~$\surface_{\bar Q}$ is
      \[
       g = \frac{|Q_1| - |Q_0| - b - p + 2}{2},
      \]
      where~$b$ is the number of boundary components (see above for a way to compute~$b$) and~$p$ the number of punctures (\ie infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$).
      Indeed, $2g = 2-\chi(\surface_{\bar Q})$, where~$\chi(\surface_{\bar Q})$ is the Euler characteristic of~$\surface_{\bar Q}$.
      The dissection~$\dissection$ provides a cellular decomposition of~$\surface_{\bar Q}$ on which we fill the boundary components and the punctures with disks.  
      The number of faces of this cellular decomposition is~$b+p$, its number of edges is~$|Q_0|$ and its number of vertices is the number of finite straight walks in~$\bar Q$, which is equal to~$2|Q_0| - |Q_1|$.
      The above formula for~$g$ follows.
\end{enumerate}
\end{remark}

\begin{proposition}
\label{prop:dualityKoszul2}
For any gentle bound quiver~$\bar Q$ with Koszul dual~$\bar Q\koszul$, the surfaces~$\surface_{\bar Q}$ and~$\surface_{\bar Q\koszul}$ coincide, but~$\dissection_{\bar Q\koszul} = {\dissection\dual_{\bar Q}}$ and~${\dissection\dual_{\bar Q\koszul}} = \dissection_{\bar Q}$.
\end{proposition}

\begin{remark}
\label{rem:comparisonWithOPS}
The construction of the surface given in Definition \ref{def:surfaceQuiver} yields the same surface as the one constructed in \cite{OpperPlamondonSchroll}.
A notable difference is that in \cite{OpperPlamondonSchroll}, only~$\dissection_{\bar Q}$ is given, and~$\dissection\dual_{\bar Q}$ is deduced (it is called the dual lamination).
Another minor difference is that the construction in \cite{OpperPlamondonSchroll} is written only for gentle algebras, while here it is generalized to locally gentle algebras.

@@

\begin{theorem}
\label{thm:complexesCoincide}
The non-kissing and non-crossing complexes are isomorphic:
\begin{itemize}
\item for any locally gentle bound quiver~$\bar Q$, the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection\dual_{\bar Q}}]$,% therefore to the accordion complex~$\AC[\dissection_{\bar Q}]$ and to the slalom complex~$\SC[{\dissection\dual_{\bar Q}}]$.
\item for any pair of dual cellular dissections~$\dissection, \dissection\dual$ of an oriented surface, the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.

@@

\begin{remark}
Note that our construction of the surface~$\surface_{\bar Q}$ implies that the vertices of~$\bar Q$ are naturally identified with the intersection points of the dual edges~$a \in \dissection_{\bar Q}$ and~$a\dual \in {\dissection\dual_{\bar Q}}$.
If we force all $\dissection$-accordions (or $\dissection\dual$-slaloms) to cross the edges of~$\dissection$ and~$\dissection\dual$ only at these intersection points, then a $\dissection$-accordion (or $\dissection\dual$-slalom) is really seen as a walk on~$\bar Q$.
\end{remark}

@@

\begin{lemma}
\label{lemm:facesHaveMaximalElement}
For any finite non-kissing face~$F$ of~$\NKC$ and any arrow~$\alpha\in Q_1\blossom$, the set~$F_\alpha$ is either empty or admits a maximal element for the countercurrent order~$\prec_\alpha$.
\end{lemma}
\begin{proof}
 We provide two proofs of the lemma.  One is combinatorial and applies to~$\NKC$, and the other is geometric and applies to~$\NCC[\dissection_{\bar Q}, {\dissection\dual_{\bar Q}}]$, using the isomorphism of Theorem \ref{thm:complexesCoincide} and Remark \ref{rem:countercurrentOrderOnSurface}.

@@

 \para{Geometric proof}
 View~$F$ as a face of the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection\dual_{\bar Q}}]$.
 The arrow~$\alpha$ then corresponds to an angle in~$\dissection$ at a vertex~$v$.

@@

\begin{proposition}
\label{prop:facetsAreFinite}
Any face of~$\NKC$ and~$\NCC[\dissection_{\bar Q}, {\dissection\dual_{\bar Q}}]$ is finite, thus contained in a finite maximal face.
\end{proposition}

Even though Theorem \ref{thm:complexesCoincide} ensures that~$\NKC$ and~$\NCC[\dissection_{\bar Q}, {\dissection\dual_{\bar Q}}]$ coincide, we provide two proofs.
One is combinatorial and applies to~$\NKC$, and the other is geometric and applies to~$\NCC[\dissection_{\bar Q}, {\dissection\dual_{\bar Q}}]$.
The proof requires the following intermediate results.

@@

 \para{Geometric proof}
 Let~$F$ be a face of the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection\dual_{\bar Q}}]$.
 Its elements are pairwise compatible~$\dissection$-accordions on the surface~$\surface_{\bar Q}$.
 Thus~$F$ can be viewed as a partial tagged triangulation of~$\surface_{\bar Q}$, by letting the~$\dissection$-accordions circling a puncture in (say) clockwise order be seen as ``notched'' arcs to said puncture, and the ones circling it in counter-clockwise orientation be ``un-notched'' arcs.
 It is known (see, for instance, \cite{FominShapiroThurston}) that any such partial tagged triangulation can be embedded into a full triangulation, and that triangulations contain a finite number of arcs.
 Thus~$F$ is finite, and is contained in a finite maximal face.
\end{proof}

@@

\begin{proof}
We provide two proofs of the lemma.
One is combinatorial and applies to~$\NKC$, and the other is geometric and applies to~$\NCC[\dissection_{\bar Q}, {\dissection\dual_{\bar Q}}]$.

>>>>

\begin{definition}%[\cite{PaluPilaudPlamondon}]
\label{def:blossomingQuiver}
A locally gentle bound quiver~$\bar Q$ is \defn{complete} if any vertex~$v \in Q_0$ is incident to either one ($v$ is a leaf) or four arrows ($v$ is an internal vertex).
The \defn{pruned subquiver} of a quiver~$\bar Q$ is the locally gentle quiver obtained by deleting all leaves of~$\bar Q$ (degree one vertices) and their incident arrows.
The \defn{blossoming quiver} of a locally gentle bound quiver~$\bar Q$ is the complete locally gentle bound quiver~$\bar Q\blossom$ whose pruned subquiver is~$\bar Q$.
The vertices of~$Q\blossom_0 \ssm Q_0$ are called \defn{blossom vertices}, and the arrows in~$Q\blossom_1 \ssm Q_1$ are called \defn{blossom arrows}.
\end{definition}

In other words, $\bar Q\blossom$ is obtained from~$\bar Q$ by adding blossom arrow and blossom vertices at each incomplete vertex of~$\bar Q$ and by completing~$I$ accordingly.
Examples are represented on \fref{fig:quivers}.

@@

\begin{definition}
\label{def:string}
Let~$\bar Q = (Q,I)$ be a locally gentle bound quiver.
A \defn{finite string} in~$\bar Q$ is a word of the form
\(
\rho = \alpha_1^{\varepsilon_1}\alpha_2^{\varepsilon_2}\cdots \alpha_\ell^{\varepsilon_\ell},
\)
where
	\begin{enumerate}[(i)]
	\item $\alpha_i \in Q_1$ and~$\varepsilon_i \in \{-1,1\}$ for all~$i \in [\ell]$,
	\item $t(\alpha_i^{\varepsilon_i}) = s(\alpha_{i+1}^{\varepsilon_{i+1}})$ for all~$i \in [\ell-1]$,
	\item there is no path~$\pi \in I$ such that~$\pi$ or~$\pi^{-1}$ appears as a factor of~$\rho$, and
	\item $\rho$ is reduced, in the sense that no factor~$\alpha\alpha^{-1}$ or~$\alpha^{-1}\alpha$ appears in~$\rho$, for~$\alpha \in Q_1$.
	\end{enumerate}
The integer~$\ell$ is called the \defn{length} of the string~$\rho$.
We let~$s(\rho) \eqdef s(\alpha_1^{\varepsilon_1})$ and~$t(\rho) \eqdef t(\alpha_\ell^{\varepsilon_\ell})$ denote the source and target of~$\rho$.
For each vertex~$v \in Q_0$, there is also a \defn{string of length zero}, denoted by~$\varepsilon_v$, that starts and ends at~$v$.
%We denote by~$\strings(\bar Q)$ the set of strings on~$\bar Q$.
%We usually denote strings by~$\rho, \sigma, \tau$.
%We call \defn{negative simple string} a formal word of length zero of the form $-v$, where $v$ is any vertex of $Q_0$. We denote by
%\(
%\strings^\pm_{\ge -1}(\bar Q) \eqdef \strings^\pm(\bar Q) \sqcup \set{-v}{v \in Q_0}
%\)
%the set of \defn{almost positive unoriented strings}.
%Note that~$\varepsilon_v$ and~$-v$ are different almost positive strings.
%\end{enumerate}
\end{definition}

\begin{definition}
 An oriented cycle~$c$ in~$Q$ such that~$c, c^2 \notin I$ is called \defn{primitive} if it cannot be written as an $n$-th power ($n>1$) of a cycle.
\end{definition}

\begin{notation}
If~$c$ is an oriented cycle in~$Q$ such that~$c, c^2 \notin I$, we write:
\begin{itemize}
\item $c^\infty$ for the infinite word~$c c c\cdots$,
\item ${}^\infty c$ for the infinite word~$\cdots c c c$,
% \item $^{-\infty} c$ for the infinite word~$\cdots c^{-1} c^{-1} c^{-1}$,
% \item $c^{-\infty}$ for the infinite word~$c^{-1} c^{-1} c^{-1}\cdots$, and
\item $^\infty c^\infty$ for the bi-infinite word~$^\infty c c^\infty$.
\end{itemize}
We also define~${}^{-\infty}c \eqdef {}^{\infty}(c^{-1}) = (c^\infty)^{-1}$ and similarly~$c^{-\infty} \eqdef (c^{-1})^\infty = (^\infty c)^{-1}$.
\end{notation}

@@

\begin{notation}
To avoid distinguishing between finite, infinite and bi-infinite words, we denote strings as products~$\rho = \prod_{i < \ell < j} \alpha_\ell$ where~$i < j \in \Z \cup \{\pm \infty\}$ and~$\alpha_\ell \in Q_0$ for all~$i < \ell < j$.
\end{notation}

@@

\begin{definition}
A \defn{marked surface}~$\bar\surface \eqdef (\surface, M)$ is an orientable surface~$\surface$ with boundaries, together with a set~$M$ of marked points which can be on the boundary of~$\surface$ or not.
For~$V \subset \surface$,
\begin{enumerate}[(i)]
\item a \defn{$V$-arc} on~$\bar\surface$ is a curve on~$\surface$ connecting two points of~$V$ and whose interior is disjoint from~$M$ and the boundary of~$\surface$.
\item a \defn{$V$-curve} on~$\bar\surface$ is a curve on~$\surface$ which at each end either reaches a point of~$V$ or infinitely circles around and finally reaches a puncture of~$M$, and whose interior is disjoint from~$M$ and the boundary of~$\surface$.
\end{enumerate}
As usual, curves and arcs are considered up to homotopy relative to their endpoints in~$\surface \ssm M$, and curves homotopic to a boundary are not allowed.
\end{definition}

@@

Some examples of dual cellular dissections on different surfaces are represented in \fref{fig:dissections}.
%
\begin{figure}[t]
	\capstart
	\centerline{\includegraphics[scale=.7]{dissections}}
	\caption{Some pairs~$(\dissection, \dissection\dual)$ of dual cellular dissections on different surfaces. The dissection~$\dissection$ is in green while its dual dissection~$\dissection\dual$ is~in~red. The boundaries of the surfaces are shaded, and the glue symbols indicate how to identify some edges to get the desired surfaces. The first two surfaces are disks, the third is a sphere without boundary, the fourth is a cylinder, and the fifth is a torus with one boundary component.}
	\label{fig:dissections}
\end{figure}
%
Note that contrarily to the usual conventions, the dual vertex~$f\dual$ of a face~$f$ of~$\dissection$ is not always in the interior of the face~$f$.
More precisely, there are two situations:
\begin{itemize}
\item if a face~$f$ has no edge on the boundary of~$\surface$, its dual vertex~$f\dual$ lies in the interior~of~$f$ and is a puncture of~$\bar\surface$,
\item if a face~$f$ has an edge on the boundary of~$\surface$, then it has exactly one such edge and its dual vertex~$f\dual$ lies on this edge.
\end{itemize}
In fact, the second point forces the following characterization of the cellular dissections that admit a dual cellular dissection.

@@

In this section, we show that non-kissing and non-crossing complexes actually coincide.
Namely, any pair of dual cellular dissections~$(\dissection, \dissection\dual)$ defines a locally gentle bound quiver~$\bar Q_{\dissection}$ such that the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection}]$.
Conversely, any locally gentle bound quiver~$\bar Q$ gives rise to a surface~$\surface_{\bar Q}$ equipped with a pair of dual cellular dissections~$\dissection_{\bar Q}$ and~${{\dissection\dual}\!\!_{\bar Q}}$ such that the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection\dual\!\!_{\bar Q}}]$.
In fact, we show that the accordion, slalom, and non-kissing complexes can all be pictured in a unique way on the surface.
Interestingly, via the correspondence between pairs of dual cellular dissections and locally gentle bound quivers, the duality between the dissections~$\dissection$ and~$\dissection\dual$ translates into the Koszul duality between the locally gentle algebras of~$\bar Q$ and~$\bar Q\koszul$.

@@

The advantage of Definition~\ref{def:surfaceQuiver} is that it automatically endows~$\surface_{\bar Q}$ with two disjoint sets~$V_{\bar Q}$ and~${V\dual_{\bar Q}}$ of marked points and two dual cellular dissections~$\dissection_{\bar Q}$ and~${\dissection\dual\!\!_{\bar Q}}$ defined as follows.

\begin{definition}
\label{def:dissectionQuiver}
The surface~$\surface_{\bar Q}$ is endowed with
\begin{itemize}
\item the set~$V_{\bar Q}$ of points~$v(\alpha)$ for~$\alpha \in Q_1\blossom$ after the identifications given by~(ii),
\item the $V_{\bar Q}$-dissection~$\dissection_{\bar Q}$ given by all sides~$\Enrs{\alpha}$ and~$\Enrt{\alpha}$ for~$\alpha \in Q_1\blossom$ after the identifications given by~(ii).
\end{itemize}
The set~${V\dual_{\bar Q}}$ and the ${V\dual_{\bar Q}}$-dissection~${\dissection\dual\!\!_{\bar Q}}$ are defined dually.
\end{definition}

\begin{proposition}
\label{prop:dissectionsAreCellular}
Let~$\bar Q$ be a locally gentle bound quiver.
Then the dissections~$\dissection_{\bar Q}$ and~$\dissection\dual\!\!_{\bar Q}$ are cellular and dual to each other.
\end{proposition}
\begin{proof}
In the construction of~$\surface_{\bar Q}$, if two arrows~$\alpha$ and~$\beta$ of~$\bar Q \blossom$ are such that~${t(\alpha) = s(\beta)}$ and~${\alpha\beta \in I\blossom}$, then the vertices~$v(\alpha)$ and~$v(\beta)$ are identified on the surface.
Thus, any two arrows of a given path contribute the same vertex of~$V_{\bar Q}$, so~$V_{\bar Q}$ is in bijection with the set of maximal paths in~$\bar Q\blossom$.

@@

\begin{remark}
\label{rem:propertiesSurface}
The following observations are useful for the computation of examples.
\begin{enumerate}[(i)]
\item The set~$V_{\bar Q}$ has one vertex for each straight walk in~$\bar Q$ (equivalently, for each maximal path in~$\bar Q$).
      Finite straight walks yield vertices on the boundary, while infinite cyclic straight walks in~$\bar Q$ and in~$\bar Q\koszul$ yield punctures.
\item \label{item:arcs}
      The dissection~$\dissection_{\bar Q}$ has one edge for each vertex~$v \in \bar Q_0$, obtained by concatenation of the sides~$\Enrt{\alpha} = \Enrs{\beta}$ and~$\Enrt{\alpha'} = \Enrs{\beta'}$ where~$v = t(\alpha) = s(\beta) = t(\alpha') = s(\beta')$, $\alpha\beta \notin I$ and~$\alpha'\beta' \notin I$. We denote by~$\edgeof(v)$ the edge of~$\dissection_{\bar Q}$ corresponding to~$v$.
\item The dissection~$\dissection_{\bar Q}$ has one $\ell$-cell for each straight walk of length~$\ell$ in~$\bar Q\koszul$.
\item A similar remark holds dually for~${V\dual_{\bar Q}}$ and~${\dissection\dual\!\!_{\bar Q}}$.
\item The number~$p$ of punctures of~$\surface_{\bar Q}$ is the number of infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$.
\item \label{item:boundary}
      The number~$b$ of boundary components of~$\surface_{\bar Q}$ can be computed as follows.
      There are two natural perfect matchings whose vertices are the blossom vertices of~$\bar Q$: one is obtained by joining the endpoints of each finite straight walk of~$\bar Q$, and the other is obtained similarly from~$\bar Q \koszul$.
      Let~$G$ be the superposition of these two perfect matchings.
      Then the number~$b$ of boundary components of~$\surface_{\bar Q}$ is the number of connected components of~$G$.
\item The genus of the surface~$\surface_{\bar Q}$ is
      \[
       g = \frac{|Q_1| - |Q_0| - b - p + 2}{2},
      \]
      where~$b$ is the number of boundary components (see above for a way to compute~$b$) and~$p$ the number of punctures (\ie infinite straight walks in~$\bar Q$ and in~$\bar Q\koszul$).
      Indeed, $2g = 2-\chi(\surface_{\bar Q})$, where~$\chi(\surface_{\bar Q})$ is the Euler characteristic of~$\surface_{\bar Q}$.
      The dissection~$\dissection$ provides a cellular decomposition of~$\surface_{\bar Q}$ on which we fill the boundary components and the punctures with disks.  
      The number of faces of this cellular decomposition is~$b+p$, its number of edges is~$|Q_0|$ and its number of vertices is the number of finite straight walks in~$\bar Q$, which is equal to~$2|Q_0| - |Q_1|$.
      The above formula for~$g$ follows.
\end{enumerate}
\end{remark}

\begin{proposition}
\label{prop:dualityKoszul2}
For any gentle bound quiver~$\bar Q$ with Koszul dual~$\bar Q\koszul$, the surfaces~$\surface_{\bar Q}$ and~$\surface_{\bar Q\koszul}$ coincide, but~$\dissection_{\bar Q\koszul} = {\dissection\dual\!\!_{\bar Q}}$ and~${\dissection\dual\!\!_{\bar Q\koszul}} = \dissection_{\bar Q}$.
\end{proposition}

\begin{remark}
\label{rem:comparisonWithOPS}
The construction of the surface given in Definition \ref{def:surfaceQuiver} yields the same surface as the one constructed in \cite{OpperPlamondonSchroll}.
A notable difference is that in \cite{OpperPlamondonSchroll}, only~$\dissection_{\bar Q}$ is given, and~$\dissection\dual\!\!_{\bar Q}$ is deduced (it is called the dual lamination).
Another minor difference is that the construction in \cite{OpperPlamondonSchroll} is written only for gentle algebras, while here it is generalized to locally gentle algebras.

@@

\begin{theorem}
\label{thm:complexesCoincide}
The non-kissing and non-crossing complexes are isomorphic:
\begin{itemize}
\item for any locally gentle bound quiver~$\bar Q$, the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection\dual\!\!_{\bar Q}}]$,% therefore to the accordion complex~$\AC[\dissection_{\bar Q}]$ and to the slalom complex~$\SC[{\dissection\dual\!\!_{\bar Q}}]$.
\item for any pair of dual cellular dissections~$\dissection, \dissection\dual$ of an oriented surface, the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.

@@

\begin{remark}
Note that our construction of the surface~$\surface_{\bar Q}$ implies that the vertices of~$\bar Q$ are naturally identified with the intersection points of the dual edges~$a \in \dissection_{\bar Q}$ and~$a\dual \in {\dissection\dual\!\!_{\bar Q}}$.
If we force all $\dissection$-accordions (or $\dissection\dual$-slaloms) to cross the edges of~$\dissection$ and~$\dissection\dual$ only at these intersection points, then a $\dissection$-accordion (or $\dissection\dual$-slalom) is really seen as a walk on~$\bar Q$.
\end{remark}

@@

\begin{lemma}
\label{lemm:facesHaveMaximalElement}
For any finite non-kissing face~$F$ of~$\NKC$ and any arrow~$\alpha\in Q_1\blossom$, the set~$F_\alpha$ is either empty or admits a maximal element for the countercurrent order~$\prec_\alpha$.
\end{lemma}
\begin{proof}
 We provide two proofs of the lemma.  One is combinatorial and applies to~$\NKC$, and the other is geometric and applies to~$\NCC[\dissection_{\bar Q}, {\dissection\dual\!\!_{\bar Q}}]$, using the isomorphism of Theorem \ref{thm:complexesCoincide} and Remark \ref{rem:countercurrentOrderOnSurface}.

@@

 \para{Geometric proof}
 View~$F$ as a face of the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection\dual\!\!_{\bar Q}}]$.
 The arrow~$\alpha$ then corresponds to an angle in~$\dissection$ at a vertex~$v$.

@@

\begin{proposition}
\label{prop:facetsAreFinite}
Any face of~$\NKC$ and~$\NCC[\dissection_{\bar Q}, {\dissection\dual\!\!_{\bar Q}}]$ is finite, thus contained in a finite maximal face.
\end{proposition}

Even though Theorem \ref{thm:complexesCoincide} ensures that~$\NKC$ and~$\NCC[\dissection_{\bar Q}, {\dissection\dual\!\!_{\bar Q}}]$ coincide, we provide two proofs.
One is combinatorial and applies to~$\NKC$, and the other is geometric and applies to~$\NCC[\dissection_{\bar Q}, {\dissection\dual\!\!_{\bar Q}}]$.
The proof requires the following intermediate results.

@@

 \para{Geometric proof}
 Let~$F$ be a face of the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection\dual\!\!_{\bar Q}}]$.
 Its elements are pairwise compatible~$\dissection$-accordions on the surface~$\surface_{\bar Q}$.
 Thus~$F$ can be viewed as a partial tagged triangulation of~$\surface_{\bar Q}$, by letting the~$\dissection$-accordions circling a puncture in (say) clockwise order be seen as ``notched'' arcs to said puncture, and the ones circling it in counter-clockwise orientation be ``un-notched'' arcs.
 It is known (see, for instance, \cite{FominShapiroThurston}) that any such partial tagged triangulation can be embedded into a full triangulation, and that triangulations contain a finite number of arcs.
 Thus~$F$ is finite, and is contained in a finite maximal face.
\end{proof}

@@

\begin{proof}
We provide two proofs of the lemma.
One is combinatorial and applies to~$\NKC$, and the other is geometric and applies to~$\NCC[\dissection_{\bar Q}, {\dissection\dual\!\!_{\bar Q}}]$.
